---
title: "Capstone Project- Credit Card Default- Taiwanaese Company "
author: "Uzair Faruqi"
date: "`r Sys.Date()`"
output: pdf_document
---

# Credit Card Default Data- for a Taiwanese Company

In This Project I have Worked with the Taiwanese Credit card Data set.
This data set has 30000 rows and 24 columns. The objective is to estimate  the probability of default payment by credit card client using the data provided. These attributes are related to various details about a customer, his past payment information and bill statements.  The Data set if available on the github repository 
https://github.com/uzaifaru2/Credit-Card-Default  it is also available on  the UCI Machine Learning repository.

I am anyways picking up the data from the repository and hence the code should run directly 

## Synopsis

The Objective of the assignment is to  predict whether a particular customer will default payment next month or not. The result is a an extremely valuable piece of information for the bank to take decisions regarding offering credit to its customer and could  affect the bank’s revenue. 

The data set is a tricky one as it has a mix of categorical and continuous variables also it is a classic case of an imbalanced data set and hence any model developed on the entire data set would have an element of bias as defined  by the overall Accuracy which these models which many of the Machine learning Models will try and Maximize.

I would be  be trying to address this  issue using various means. 
First I will try and develop a normal logistic regression model to predict the probability default, measuring the probability at two different  benchmarks of 0.5 and 0.25 and see the overall accuracy as well as the specificity and sensitivity.

Then next I would be  developing a decision tree model using the rpart library to achieve the objective.
In the decision tree I would be utilizing these 4 methods available in the rpart package to address the issue 
1.	Under Sampling the training set
2.	Changing the prior probability
3.	Assigning  a Cost to Default
4.	Changing the weights to default vs non Default. data
As we see that when we develop a model  on the entire set , the Accuracy might seem good but the specificity or Sensitivity  gets compromised.

With the Various versions of the respective Tree models as we see that we can balance  the Objectives  in terms of selecting the models which has a Better Sensitivity or Specificity.

We will round it off by developing a Boosting model i.e a Gradient Boosting model, Multi Variate Adaptive Recursive Slime - Better Known as earth  and Adaboost and see the confusion matrix developed on the same. The Boosting  models have all been trained on the Undersampled Training set where we have made attempts to balance the data set. I have also tried to develop the model on the full data set. I would be comparing the performance of the various models or the various approaches applied.

As we see that we would have to strike e a balance between the Specificity and the Sensitivity of the models to achieve  the Objectives .


## Loading the necessary libraries and Extracting the Data 



```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}


library(tidyverse)
library(broom)
library(caret)
library(ipred)
library(lattice)
library(ggplot2)
library(caret)
library(MASS)
library(car)
library(knitr)
library(mice)
library(lattice)
library(reshape2)
library(skimr)
library(ROSE)
library(rpart)
library(rpart.plot)
library(earth)

data <- read.csv('https://raw.githubusercontent.com/uzaifaru2/Credit-Card-Default/main/credit_card_default_data.csv')





```


```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}

# Exploring the data

dim(data)

# We are having 30,000 records with around 25 columns - variables ( including a customer ID column)
# The defaults  have been encoded as 1 and 0 otherwise 


# All the variables including the customer ID seems to have been considered as integers



# Looking at the data to see if any Inconsistency 

table(data$EDUCATION)

# There are data with 5 and 6 defined for Education  which hasn't been defined in the data 
# data understanding document. We might have to group them under 4 which  is others

table(data$MARRIAGE)

# here also we have many under 0 , which has not been defined , we can put it under 3 Which is Others


#replace 0's with NAN, replace others too
data$EDUCATION[data$EDUCATION == 0] <- 4
data$EDUCATION[data$EDUCATION == 5] <- 4
data$EDUCATION[data$EDUCATION == 6] <- 4
data$MARRIAGE[data$MARRIAGE == 0] <- 3

# Checking on whether changes have been Made

table(data$EDUCATION)
table(data$MARRIAGE)





```


### About the Data

There are around 30,000 data points i.e around 30,000 unique data points 
All the data including the customer ID have been considered as integers
We would have to make the necessary corrections, converting some of the data into factors  and some of the variables might have to be one hot encoded so that we can run the models relate to the Boosting Algorithms





## Creating Derived Variables 

One of the Important Tasks of a Data scientist is to understand the data  and to work on making certain derived variables which  would be more effective as variables  then the Variables themselves

For Example the Billing Amount or the credit balance might not be so useful themselves say as compared to say the Ratio of Billing Amount to the Credit balance .

There could be many such variables that can be derived and used in the  model developed to improve on its accuracy. Much of it depends on the ingenuity of the Data Scientists and the understanding of the domain by them.

Also while  we might derive many variables there would have to be certain data engineering to be done to address anomalies etc



```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}


# Let us work out certain Derived variables 
# One can be as innovative as possible  in Deriving these variables and these variables 
# Can actually effect the performance of the model itself also.

# The Derived Variables that we can look at are
# The increase in the Billing amount in september as compared to August
# The Billing amount in a percentage of the Credit limit
# Payment to Bill AMNT Ration

data2 <-data %>% mutate( inc_billing_amnt = (BILL_AMT1- BILL_AMT2)/BILL_AMT2 , 
                         bill_amnt_Limit = BILL_AMT1/LIMIT_BAL,
                         pay_bill_amnt_ratio = PAY_AMT1/BILL_AMT2)


# Now Another Problem , what needs to be done for cases which are  Giving NA's or Inf/-inf

# One Option is to remove these variables another option might be to replace the NA's by 
# 0's  or the maximum values
# Infy can be handled by replacing the divisor ( if it was a 0 by the median value) but it will end 
# up changing a variable itself which  we should avoid.
# Depending on the prevalence of these values

# write.csv(data2, file = "data2.csv")

# We are going to replace the NA's by 0 and the Infinities by the maximum value in that column


max(data2$inc_billing_amnt[is.finite(data2$inc_billing_amnt)])

data2$inc_billing_amnt[data2$inc_billing_amnt == Inf] <- max(data2$inc_billing_amnt
                                                        [is.finite(data2$inc_billing_amnt)])
data2$inc_billing_amnt[data2$inc_billing_amnt == -Inf] <- min(data2$inc_billing_amnt
                                                          [is.finite(data2$inc_billing_amnt)])
data2$inc_billing_amnt[is.na(data2$inc_billing_amnt) ] <- 0

# managing the Infinity Values for the  pay_bill_amnt_ratio both the negative and positive Infinities


max(data2$pay_bill_amnt_ratio[is.finite(data2$pay_bill_amnt_ratio)])
min(data2$pay_bill_amnt_ratio[is.finite(data2$pay_bill_amnt_ratio)])

data2$pay_bill_amnt_ratio[data2$pay_bill_amnt_ratio == Inf] <- max(data2$pay_bill_amnt_ratio[is.finite(data2$pay_bill_amnt_ratio)])
data2$pay_bill_amnt_ratio[data2$pay_bill_amnt_ratio == -Inf] <- min(data2$pay_bill_amnt_ratio[is.finite(data2$pay_bill_amnt_ratio)])
data2$pay_bill_amnt_ratio[is.na(data2$pay_bill_amnt_ratio) ] <- 0




# Checking for NA's

sapply(data2, function(x) sum(is.na(x)))





```

## Data Engineeringa and creating the Training set and the Test Set  

Now we would be working on making the data ready for  the models being developed.
This would Involve  handling the NA's if any, Doing One Hot Encoding since I intend to use the Boosting algorithms to measure their efficacy.
Handling the Numerical data  as well as factors since  though the target variable is stored as a 1 or a 0 it is stored as integer which  we might have to convert into factors for the Algo to work properly. 



```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}

# lets remove the ID columns from the data set
data2 <- data2[ , -1]



# We see that certain variables which  should be categorical are stored as integers or 
# numeric manner
# These should be converted into factors for further analysis
# The variables which  should be factors are " Sex", " Education " , " Marriage",  
# Using them as numeric would give a false indication of their relevance for the End 
# Many of the variables like the History of the payment status can also be used as numeric
# but as numeric they are not significant as conveyed by the correlation matrix and 
# hence we are using  # these are factors


names <- c(2:4)

data2[,names] <- lapply(data2[,names] , factor)
# str(data2)
#One-Hot Encoding
# Creating dummy variables is converting a categorical variable to as many binary variables as 
#mhere are categories.
dmy <- dummyVars(" ~ .", data = data2,fullRank = T) 

train_transformed <- data.frame(predict(dmy, newdata = data2)) 
# # See the structure of the new dataset
# str(train_transformed)

#Converting the dependent variable back to categorical 
train_transformed$default_payment<-as.factor(train_transformed$default_payment)

# checking the same and the other variables 
#str(train_transformed)

#Splitting training set into two parts based on outcome: 75% and 25% 
set.seed(7)
index <- createDataPartition(train_transformed$default_payment, p = 0.75, list=FALSE, times = 1) 
trainSet <- train_transformed[ index,]
testSet <- train_transformed[-index,]






```



Now we have the Training set Ready and the Test Set, 
we will start with developing a Simple Logistics regression model  and see the Results 
and we can also observe the  Implications of an Imbalanced data



## Logistic Regression Model

Logistic regression is the First model usually attempted in any Classification  problem.
It is the Linear regression equivalent model as far as classification problem is concerned.

In short, Logistic Regression is used when the dependent variable(target) is categorical. For example:

To predict whether an email is spam (1) or not spam (0)
Whether the tumor is malignant (1) or not (0)

It is named as ‘Logistic Regression’, because it’s underlying technique is quite the same as Linear Regression. There are structural differences in how linear and logistic regression operate. Therefore, linear regression isn’t suitable to be used for classification problems.

Its name is derived from one of the core function behind its implementation called the logistic function or the sigmoid function. It’s an S-shaped curve that can take any real-valued number and map it into a value between 0 and 1, but never exactly at those limits.



```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}
log.model <- glm(default_payment ~., data = trainSet, family = binomial)

summary(log.model)


```



### Summarising From The Model 

By Creating the Model and Looking at the Summary of the Model we can see which of the variables are  important , i.e the ones with the stars

- Bill Amount to the Limit Ratio
- INcrease in Billing Amount
- Payment Amount 2 - Amount Paid 2 months back
- Payment Amount 1 - Payment made 1 Month Back
- Bill Amount last Month
- Bill Amount 2 months back
- Marriage
- Education 
- Increase in Billing Amount as compared to previous month
- Increase in Billing Amount as a Ratio

These variables  are important as far as creating and running a model is concerned.

I had run the model on these set of Important variables as well as the Model with all the variables . There wasnt any noticeable difference in the output and hence sticking with all the variables for this project. 



```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}
# Two of the derived variables etc

log.predictions <- predict(log.model, testSet, type="response")

log.predictions.rd <- ifelse(log.predictions > 0.5, 1, 0)
log.predictions.rd <- as.factor(log.predictions.rd)



con_matrix_log_0.5 <- confusionMatrix(log.predictions.rd,testSet$default_payment)

# With This Type of Mode The Sensitivity Seems to be High but the  Precision seems to be lower.
# Accuracy is around 80 % but that could be the case , but we could have got an accuracy of around 
# 75 % by predicting non Defaults
# Let us see how we can change things , by Changing the default probability from 0.5 to say 0.25

log.predictions.rd1 <- ifelse(log.predictions > 0.25, 1, 0)
log.predictions.rd1 <- as.factor(log.predictions.rd1)




con_matrix_log_0.25 <- confusionMatrix(log.predictions.rd1,testSet$default_payment)




# Extracting the Accuracy , Sensitivity and Specificity from the Confusion Matrix Created

log.accuracy.0.5 <- con_matrix_log_0.5$overall[["Accuracy"]]
log.accuracy.0.25 <- con_matrix_log_0.25$overall[["Accuracy"]]
log.specificity.0.5 <- con_matrix_log_0.5$byClass[["Specificity"]]
log.specificity.0.25 <- con_matrix_log_0.25$byClass[["Specificity"]]
log.sensitivity.0.5 <- con_matrix_log_0.5$byClass[["Sensitivity"]]
log.sensitivity.0.25 <- con_matrix_log_0.25$byClass[["Sensitivity"]]


```



### The Output of the Logistic Regression

The Accuracy obtained is around 80.6 %  and the Specificity 23.6 %  for the cut off  @ 0.4

The Accuracy obtained is around 75.6 %  and the Specificity increases to 54.8 %  for the cut off  @ 0.4

Here for our case the the Positive is non default and the negative is default ( Not Referring to the 1's and 0s')

Hence  the specificity measures  our efficiency in terms of identifying out of the total number of defaulters how well we have been able to identify them.

We would be evaluation various measures to do that specifically using various measures as given below 






# Tree Based Models

Now we would be working with some Tree based Models to see how they perform 

Sensitivity  and specificity   both increases  depending on the cut off chosen , accuracy Decreases,  it might all depend on the priority of the organization 

These are the typical characteristics of an imbalanced data set

We will now go through  a set of measures designed to address these type of Imbalanced data 

These Measures are 
-Undersampling ( Since the Data set is sufficiently large we can as of now avoid Over sampling)
-The Undersampling would be done only on the Training Set
-Changing the Prior Probabilities
-Including a Loss Matrix

These are some of the methods that can be utilised to address this problem of imbalanced data  there are many other methods also.

Creating an Undersample Data set from the Training Set  We would be using the Rose Library of R to do this and to develop the  respective Machine learning model.



### First Let us Look at the Proportions of the Default in the Training Set



```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}



table(trainSet$default_payment)
prop.table(table(trainSet$default_payment))

```

we See that the Proportion of defaults is around 22 % , which is a Unbalanced Dataset
we will try and Create an UNder sample Data set where the Proportion of defaults and the Proportion  of Non Defaults are equal  and we would be training  our models on this Balanced Data set and  the testing would be done on the Test Data set Developed


```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}
train_under_sample <- ovun.sample(default_payment ~ ., data = trainSet, 
                                  method = "under", N = 9954, seed = 1)$data
table(train_under_sample$default_payment)

```

### Here we are building the models using the undersample data set


```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}

# Now Lets  start Building  Tree Based Models and observe whether we are able to improve on 
# The Accuracy or the other measures like Specificity

tree_undersample <- rpart(default_payment ~ ., method = "class",
                          data =  train_under_sample, control = rpart.control(cp = 0.001))

# Now Let us Plot the Tree's

#plot(tree_undersample, uniform = TRUE)

# Add labels to the decision tree
# text(tree_undersample)

# I am avoiding PLotting the Trees , since the Image isn't coming out 
# properly in the RMD generated File , 
# it Can be seen properly in a normal R Studio Console


```

### Pruning the Trees 

We are now pruning the trees to have least error and to avoid over fitting


```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}

# Now Lets  Prune the tree for the Model tree_undersample

# Plotting the cross-validated error rate as a function of the complexity parameter
plotcp(tree_undersample)

# Using  printcp() to identify for which complexity parameter 
# the cross-validated error rate is minimized.

printcp(tree_undersample)

# Creating  an index for of the row with the minimum xerror
index0 <- which.min(tree_undersample$cptable[ , "xerror"])

# Creating tree_min
tree_min <- tree_undersample$cptable[index0, "CP"]

#  Prune the tree using tree_min
ptree_undersample <- prune(tree_undersample, cp = tree_min)

```

### Changing the Prior Probabilities

in the rpart package the prior probabilities by default are taken as per their proportions , we can sort of trick R by modifying the same in the parms function to say 60/40 


```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}

tree_prior <- rpart(default_payment ~ ., method = "class",parms = list(prior=c(0.6, 0.4)),
                    control = rpart.control(cp = 0.001), 
                    data = trainSet)

# Plot the decision tree
# plot(tree_prior, uniform = TRUE)

# Add labels to the decision tree
# text(tree_prior)

```

## Pruning the Tree 

Pruning the Tree for the Model with the Changed Prio Probabilities



```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}

# As the Trees get more complex, they are not clearly visible( The Nodes)

# Pruning the Trees with Changed Prior Probabilities

	# Plot the cross-validated error rate as a function of the complexity parameter
 plotcp(tree_prior)

# Use printcp() to identify for which complexity parameter the 
# cross-validated error rate is minimized.
printcp(tree_prior)
	
# Create an index for of the row with the minimum xerror
index <- which.min(tree_prior$cptable[ , "xerror"])

# Create tree_min
tree_min <- tree_prior$cptable[index, "CP"]

#  Prune the tree using tree_min
ptree_prior <- prune(tree_prior, cp = tree_min)

# Use prp() to plot 

prp(ptree_prior)

```


## Tree with a Loss Matrix Included

We are now developing a Tree by including a loss matrix where we are penalizing a Default 10 Times more as compared to Non Default.
An Arbitrary Number chosen. The Company can choose any other number depicting the actual loss that they might face in case of a default.


```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}


tree_loss_matrix <- rpart(default_payment ~ ., method = "class",
                          parms = list(loss = matrix(c(0, 10, 1, 0), ncol=2)),
                          control = rpart.control(cp = 0.001), data =  trainSet)

# Plot the decision tree
#plot(tree_loss_matrix, uniform = TRUE)

# Add labels to the decision tree
#text(tree_loss_matrix)
```

## Pruning the Tree

Pruning the Tree for the Model developed by including a Loss Matrix

```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}


# Pruning the tree with the loss matrix

set.seed(345)
tree_loss_matrix  <- rpart(default_payment ~ ., method = "class", data = trainSet,
                           parms = list(loss=matrix(c(0, 10, 1, 0), ncol = 2)),
                           control = rpart.control(cp = 0.001))

# Plot the cross-validated error rate as a function of the complexity parameter
plotcp(tree_loss_matrix)

printcp(tree_loss_matrix)

# Create an index for of the row with the minimum xerror
index2 <- which.min(tree_loss_matrix$cptable[ , "xerror"])

# Create tree_min
tree_min2 <- tree_loss_matrix$cptable[index2, "CP"]
# Prune the tree using cp = 0.0012788
ptree_loss_matrix <- prune(tree_loss_matrix, cp = tree_min2)

# Use prp() and argument extra = 1 to plot the pruned tree
prp(ptree_loss_matrix, extra = 1)


```

## Tree with Weights

Here we are developing Tree model by Assigning different weights to default vs Non Default 
 
```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}


# Pruning the Tree for the case with Weights 
# This vector contains weights of 1 for the non-defaults in the training set, 
# and weights of 3 for defaults in the training sets. 
# By specifying higher weights for default, the model will assign higher 
# importance to classifying defaults correctly.

# Creating a Vector of weights

case_weights <- ifelse(trainSet$default_payment == 1, 3,1 )

head(case_weights)
str(case_weights)

head(trainSet$default_payment)
set.seed(345)
tree_weights <- rpart(default_payment ~ ., method = "class",
                      data = trainSet, weights = case_weights,
                      control = rpart.control(minsplit = 5, minbucket = 2, cp = 0.001))

# Plot the cross-validated error rate for a changing cp
plotcp(tree_weights)

# Create an index for of the row with the minimum xerror
index <- which.min(tree_weights$cp[ , "xerror"])

# Create tree_min
tree_min <- tree_weights$cp[index, "CP"]

#  Prune the tree using tree_min
ptree_weights <- prune(tree_weights, tree_min)

# Plot the pruned tree using the rpart.plot()-package
prp(ptree_weights, extra = 1)

```
## Testing of the Different Tree Based Models

Now we would be predicting the probabilities of defaults on the different models by using the Predict Function and by Printing out the Confusion Matrixes



```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}

# Now Lets Predict on the test set using the different models that we have developed and 
#Compute the Confusion Matrices for the respective models

# Make predictions for each of the pruned trees using the test set.
pred_undersample <- predict(ptree_undersample, newdata = testSet,  type = "class")
pred_prior <- predict(ptree_prior, newdata = testSet,  type = "class")
pred_loss_matrix <- predict(ptree_loss_matrix, newdata = testSet,  type = "class")
pred_weights <- predict(ptree_weights, newdata = testSet,  type = "class")

# Now Lets Create the Confusion Matrices of the Resulting Models and observe the Accuracies 
#and the Specificity/Sensitivity Obtained

pred_undersample <- as.factor(pred_undersample)
pred_prior <- as.factor(pred_prior)
pred_loss_matrix <- as.factor(pred_loss_matrix)
pred_weights <- as.factor(pred_weights)




con_matrix_undersample <- confusionMatrix(pred_undersample,testSet$default_payment)
con_matrix_prior <- confusionMatrix(pred_prior,testSet$default_payment)
con_matrix_loss_matrix <- confusionMatrix(pred_loss_matrix,testSet$default_payment)
con_matrix_weights<- confusionMatrix(pred_weights,testSet$default_payment)

# Extracting the Accuracy, Specificity and the Sensitivity

ptree_undersample.accuracy <- con_matrix_undersample$overall[["Accuracy"]]
ptree_undersample.specificity <- con_matrix_undersample$byClass[["Specificity"]]
ptree_undersample.sensitivity <- con_matrix_undersample$byClass[["Sensitivity"]]

ptree_prior.accuracy <- con_matrix_prior$overall[["Accuracy"]]
ptree_prior.specificity <- con_matrix_prior$byClass[["Specificity"]]
ptree_prior.sensitivity <- con_matrix_prior$byClass[["Sensitivity"]]

ptree_loss_matrix.accuracy <- con_matrix_loss_matrix$overall[["Accuracy"]]
ptree_loss_matrix.specificity <- con_matrix_loss_matrix$byClass[["Specificity"]]
ptree_loss_matrix.sensitivity <- con_matrix_loss_matrix$byClass[["Sensitivity"]]

ptree_weights.accuracy <- con_matrix_weights$overall[["Accuracy"]]
ptree_weights.specificity <- con_matrix_weights$byClass[["Specificity"]]
ptree_weights.sensitivity <- con_matrix_weights$byClass[["Sensitivity"]]

## All the Accuracies, Specificity and Sensitivity has been stored in 
# separate variables since I Intend to compile these and print out a summarized sheet

```

# Boosting Algorithms


Unlike many ML models which focus on high quality prediction done by a single model, boosting algorithms seek to improve the prediction power by training a sequence of weak models, each compensating the weaknesses of its predecessors.

boosting is a generic algorithm rather than a specific model. Boosting needs you to specify a weak model (e.g. regression, shallow decision trees, etc) and then improves it.

We would be using two Boosting Algoriths

## MARS

Multivariate Adaptive  Regression Splines

Multivariate Adaptive Regression Splines, or MARS, is an algorithm for complex non-linear regression problems.

The algorithm involves finding a set of simple linear functions that in aggregate result in the best predictive performance. In this way, MARS is a type of ensemble of simple linear functions and can achieve good performance on challenging regression problems with many input variables and complex non-linear relationships.


##  AdaBoost

AdaBoost is a specific Boosting algorithm developed for classification problems (also called discrete AdaBoost). The weakness is identified by the weak estimator’s error rate.

In each iteration, AdaBoost identifies miss-classified data points, increasing their weights (and decrease the weights of correct points, in a sense) so that the next classifier will pay extra attention to get them right.


### Creating the Training Set and the Test Set for the Boosting Algorithms 

We would be creating the models for both the undersample data as well as the Full data set


```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}

# Creating the Training Set and the Test Set
# Here we will be scaling the Data to see the performance of the models with Scaled Data
# We Try and Build These models on the Entire set as well as the Undersampled data set 
# and Observe The accuracies and the Specificity/Sensitivity Achieved 

# Training the Complete Data set

trainSet1 <- trainSet %>% mutate(default_payment = 
                                   factor(ifelse(default_payment == "1", "Yes", "No")))
preProcess_range_model <- preProcess(trainSet1, method='range')
trainSet1 <- predict(preProcess_range_model, newdata = trainSet1)

train_under_sample1 <- train_under_sample %>% mutate(default_payment = 
                                    factor(ifelse(default_payment == "1", "Yes", "No")))

preProcess_range_model1 <- preProcess(train_under_sample1, method = 'range')
train_under_sample1 <- predict(preProcess_range_model1, newdata = train_under_sample1)

# Preparing the test Set

testSet1 <- testSet %>% mutate(default_payment = 
                          factor(ifelse(default_payment == "1", "Yes", "No")))

# Scaling the Test Set for the Full Model

testSetFull <- predict(preProcess_range_model, newdata = testSet1)
testSet_under_sample1 <- predict(preProcess_range_model1, newdata = testSet1)
```


```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}
# Developing a MARS Model- Multivariate Adaptive Regression Spline  to
# Predict the defaults on the Full data , in the later part
# we would be developing and Testing  it on the undersample Data
# We would be developing this model using both Tune Length and Tune Grid on the 
# Undersampled Training Set
# we would be Observing the Accuracy and the Sensitivity as we as the Specificity to 
# understand the performance of the model
# Define the training control


fitControl <- trainControl(
  method = 'cv',                   # k-fold cross validation
  number = 5,                      # number of folds
  savePredictions = 'final',       # saves predictions for optimal tuning parameter
  classProbs = TRUE,                  # should class probabilities be returned
  summaryFunction=twoClassSummary  # results summary function
) 


# Step 1: Tune hyper parameters by setting tuneLength
set.seed(100)
model_mars1 <- train(default_payment ~ ., data=trainSet1 , method='earth',
                     tuneLength = 5, metric='ROC', trControl = fitControl)
# model_mars1

# Step 2: Predict on testData and Compute the confusion matrix

predictedMars1 <- predict(model_mars1, testSetFull)

con_matrix_Mars <- confusionMatrix(predictedMars1,testSetFull$default_payment)

Mars.accuracy <- con_matrix_Mars$overall[["Accuracy"]]
Mars.specificity <- con_matrix_Mars$byClass[["Specificity"]]
Mars.sensitivity <- con_matrix_Mars$byClass[["Sensitivity"]]

# Doing the predictions Using a Tune Grid

# Step 1: Define the tuneGrid
marsGrid <-  expand.grid(nprune = c(2, 4, 6, 8, 10), 
                         degree = c(1, 2, 3))

# Step 2: Tune hyper parameters by setting tuneGrid
set.seed(100)
model_mars2 = train(default_payment ~ ., data=trainSet1 , method='earth', 
                    metric='ROC', tuneGrid = marsGrid, trControl = fitControl)
# model_mars2



# Step 3: Predict on testData and Compute the confusion matrix
predictedMars2 <- predict(model_mars2, testSetFull)

con_Matrix_Mars_Tuned <- confusionMatrix(predictedMars2,testSetFull$default_payment)



Mars_Tuned.accuracy <- con_Matrix_Mars_Tuned$overall[["Accuracy"]]
Mars_Tuned.specificity <- con_Matrix_Mars_Tuned$byClass[["Specificity"]]
Mars_Tuned.sensitivity <- con_Matrix_Mars_Tuned$byClass[["Sensitivity"]]


```
### Devloping the MARS model on the undersampled data


```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}
#  Developing a MARS Model- Multivariate Adaptive Regression Spline  to Predict the defaults 
# on the under sampled Data
# We would be developing this model using both Tune Length and Tune Grid on the 
# Undersampled Training Set
# we would be Observing the Accuracy and the Sensitivity as we as the Specificity to 
# understand the performance of the model
# Define the training control


fitControl <- trainControl(
  method = 'cv',                   # k-fold cross validation
  number = 5,                      # number of folds
  savePredictions = 'final',       # saves predictions for optimal tuning parameter
  classProbs = TRUE,                  # should class probabilities be returned
  summaryFunction=twoClassSummary  # results summary function
) 


# Step 1: Tune hyper parameters by setting tuneLength
set.seed(100)
model_mars1_under_sample  <- train(default_payment ~ ., data=train_under_sample1 ,
                    method='earth', tuneLength = 5, metric='ROC', trControl = fitControl)
# model_mars1_under_sample

# Step 2: Predict on testData and Compute the confusion matrix

predictedMars1_under_sample <- predict(model_mars1_under_sample, testSet_under_sample1)

con_Matrix_Mars_Under_sample <- confusionMatrix(predictedMars1_under_sample, 
                                                testSet_under_sample1$default_payment)





Mars_Under_Sample.accuracy <- con_Matrix_Mars_Under_sample$overall[["Accuracy"]]
Mars_Under_Sample.specificity <- con_Matrix_Mars_Under_sample$byClass[["Specificity"]]
Mars_Under_Sample.sensitivity <-con_Matrix_Mars_Under_sample$byClass[["Sensitivity"]]

# Doing the predictions Using a Tune Grid

# Step 1: Define the tuneGrid
marsGrid <-  expand.grid(nprune = c(2, 4, 6, 8, 10), 
                         degree = c(1, 2, 3))

# Step 2: Tune hyper parameters by setting tuneGrid
set.seed(100)
model_mars2_under_sample = train(default_payment ~ ., data=train_under_sample1 , 
                method='earth', metric='ROC', tuneGrid = marsGrid, trControl = fitControl)

# model_mars2_under_sample



# Step 3: Predict on testData and Compute the confusion matrix
predictedMars2_under_sample <- predict(model_mars2_under_sample, testSet_under_sample1)

con_Matrix_Mars_Tuned_Under_sample <-confusionMatrix(predictedMars2_under_sample, 
                                                testSet_under_sample1$default_payment)


Mars_Tuned_Under_Sample.accuracy <- con_Matrix_Mars_Tuned_Under_sample$overall[["Accuracy"]]
Mars_Tuned_Under_Sample.specificity <- con_Matrix_Mars_Tuned_Under_sample$byClass[["Specificity"]]
Mars_Tuned_Under_Sample.sensitivity <-con_Matrix_Mars_Tuned_Under_sample$byClass[["Sensitivity"]]
```


### Creating the AdaBoost Model 

```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}


set.seed(100)

# Train the model using adaboost
model_adaboost1 <- train(default_payment ~ ., data=trainSet1, method='adaboost',
                         tuneLength=2, trControl = fitControl)
# model_adaboost1

predicted_adaboost1 <- predict(model_adaboost1, testSetFull)
con_Matrix_adaboost <-confusionMatrix(predicted_adaboost1, testSetFull$default_payment)

Adaboost.accuracy <- con_Matrix_adaboost$overall[["Accuracy"]]
Adaboost.specificity <- con_Matrix_adaboost$byClass[["Specificity"]]
Adaboost.sensitivity <-con_Matrix_adaboost$byClass[["Sensitivity"]]

```

## Creating and testng the adaboost Model on the Under Sampled Data 


```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}



set.seed(100)

# Train the model using adaboost
model_adaboost1_under_sample = train(default_payment ~ ., data=train_under_sample1, 
                              method='adaboost', tuneLength=2, trControl = fitControl)

# model_adaboost1_under_sample

predicted_adaboost1_under_sample <- predict(model_adaboost1_under_sample, testSet_under_sample1)
confusionMatrix(predicted_adaboost1_under_sample, testSet_under_sample1$default_payment)

con_Matrix_adaboost_under_sample <-confusionMatrix(predicted_adaboost1_under_sample, testSet_under_sample1$default_payment)

Adaboost_under_sample.accuracy <- con_Matrix_adaboost_under_sample$overall[["Accuracy"]]
Adaboost_under_sample.specificity <- con_Matrix_adaboost_under_sample$byClass[["Specificity"]]
Adaboost_under_sample.sensitivity <-con_Matrix_adaboost_under_sample$byClass[["Sensitivity"]]

```

## Compiling the Results of All the Models 


Comparison of the Performance of the different Models in terms of the Overall Accuracy, Sensitivity and  Specificity 

```{r echo= TRUE,warning = FALSE,message = FALSE, results='asis'}


The_Models <- c('Logistic_Regression_Cut_Off_0.5', 'Logistic_Regression_Cut_Off_0.25',
'Pruned_Tree_Under_Sample','Pruned_Tree_Prior_Probability',
'Pruned_Tree_Loss_Matrix', 'Pruned_Tree_Weights', 
'MARS', 'MARS_Tuned',
'MARs_Under_Sample', 'MARS_Under_Sample',
'Adaboost','Adaboost_Under_Sample')

Accuracy_models <- c(log.accuracy.0.5 , log.accuracy.0.25 ,ptree_undersample.accuracy , 
ptree_prior.accuracy , ptree_loss_matrix.accuracy, ptree_weights.accuracy , Mars.accuracy ,
Mars_Tuned.accuracy, Mars_Under_Sample.accuracy, Mars_Tuned_Under_Sample.accuracy, 
Adaboost.accuracy,Adaboost_under_sample.accuracy )


Specificity_models <- c(log.specificity.0.5 , log.specificity.0.25 ,ptree_undersample.specificity , 
ptree_prior.specificity , ptree_loss_matrix.specificity, ptree_weights.specificity , Mars.specificity ,
Mars_Tuned.specificity, Mars_Under_Sample.specificity, Mars_Tuned_Under_Sample.specificity, 
Adaboost.specificity,Adaboost_under_sample.specificity )


Sensitivity_models <- c(log.sensitivity.0.5 , log.sensitivity.0.25 ,ptree_undersample.sensitivity , 
ptree_prior.sensitivity , ptree_loss_matrix.sensitivity, ptree_weights.sensitivity , Mars.sensitivity ,
Mars_Tuned.sensitivity, Mars_Under_Sample.sensitivity, Mars_Tuned_Under_Sample.sensitivity, 
Adaboost.sensitivity,Adaboost_under_sample.sensitivity )

Results <- data.frame(The_Models, Accuracy_models,Sensitivity_models ,Specificity_models)



```

## Giving  The Summary of all the Modeld developed for easy comparison 


```{r tables}
kable(Results, col.names =  c("The Models", "Accuracy ", " Sensitivity", " Specificity"),
      align = "cc", caption = "The Summary of All the Models Developed.")
```





# Conclusions




Here  in the Models developed the POsitive is considered as Non Defaults Denoted by "0" and Negative as Default denoted by "1".

We  are here more Interested in the Defaults which are denoted by 1's

Now in such a scenario as mentioned earlier  the Sensitivity Denotes the Power of the Model , i.e the accuracy wrt the Predictions made, and Specificity denotes that out of the existing defaults , how accurately has the model been able to predict the defaults.

From the Banks point of view Specificity is more important here.
In Such a Scenario the  the Prune Tree Loss matrix and the Pruned Tree Under Sample gives better results
Pruned Tree has the Best SPecificity but the accuracy is heavily compromised in order to achieve this.

The Under Sample Models of the Boosting Algorithms MARS and Adaboost giver better Specificity as compared to the Other Models while somewhat maintaining the Accuracies.

This is an unbalanced data set and these type of problems would be  faced by the data Scientists designing the ML model to predict the Default.

The Reason for trying out so many model was to check all the options from a basic model to the Boosting algorithms, From Scaling the data to not scaling it and see what are the results  whether there exists any significant  difference between the output of a basic model to a very sophisticated model. There isnt any drastic difference , the difference is minor. 

If the Power of the Model being a criteria where the option is that out of the predictions made , the predictions should be most accurate then MARS seems to be doing a Good Job. But if we want to identify the Maximum of the Defaulters than the models using an undersample data set for training does a decent job. 




Here the company might have to define its cost ,  Of not  predicting a would be defaulter correctly and what sort of default rate is sort of acceptable . Based on that Insight the cost component can be Included in the model to give appropriate results which would be useful  to the Organization which  is the credit card company.

Based on  the cost defined  an appropriate model can be designed which  would increase the relevant metric of interest to the organization , a Trade off might have to be decided upon. 







